---
title: "ENVS-193DS_homework-03"
author: "Niki van der Poel"
format: pdf
---

# Set up

```{r}
#| message: FALSE
#| label: Set up

library(tidyverse)
library(here)
library(janitor)
library(readxl)

kelp <- read_csv("data/temp-kelp.csv")
my_data <- read_xlsx("data/Project 193DS.xlsx")
```

# Problems

# Problem 1. Giant kelp fronds

## a. An appropriate test

An appropriate test to determine the strength of the relationship between temperature and giant kelp frond elongation rate would be to measure the correlation. You can measure stregth of correlation through the following tests: Pearson's product-moment correlation, Spearman's rank correlation.

## b. Create a visualization

```{r}
#| label: creating a visualization

# base layer ggplot
ggplot(data = kelp,
       # x axis = temperature
       # y axis = kelp_elong
       mapping = aes( x = temp_c,
                      y = kelp_elong)
       ) +
  # adding geometry - points, scatterplot
  geom_point(size = 2,
             # changing color from default
             color = "skyblue") +
  # changing theme from default
  theme_classic() +
  # adding axes labels
  labs( x = "Temperature (Celsius)",
        y = "Giant Kelp Frond Elongation rate (cm/day)")



```


## c. Check your assumptions and run your test

### Checking Assumptions

```{r}
#| label: checking assumptions

# creating new object
# linear regression model of kelp
# (response ~ predictor)
kelp_lm <- lm( kelp_elong ~ temp_c,
    data = kelp
    )
#displaying linear model
# putting side by side
par(mfrow = c(2,2)) 
plot(kelp_lm)

```

For assumptions I wanted to check the homoscedasticity, normal distribution, and outliers of the model using a linear regression model. The outputs demonstrate random distribuions for Residuals vs Fitted and Scale-Location indicating homoscedasticity; points that, in general, match the qq line in the QQ plot, indicating normal distribution; and not really any outliers in the Residuals vs Leverage plot as the points lay inside of the dotted lines. 

### Running Correlation test

```{r}
#| label: running correlation test

# running pearsons coefficient
# data$predictor, data$response
cor.test(kelp$temp_c, kelp$kelp_elong,
         method = "pearson")


```


## d. Results communication

I used a Pearson's Correlation test in order to determine the strength of the relationship between temperature and giant kelp frond elongation rate. I found a moderate negative relationship between temperature and giant kelp frond elongation rate (Pearson's r = -0.69, t(30) = -5.2, p < 0.001, $\alpha$ = 0.05).

## e. Test implications

We found a moderate negative (Pearson's r = -0.69) relationship between temperature and giant kelp frond elongation rate. Here, as temperature increases, giant kelp frond elongation rate decreases. 

## f. Double check your own work

```{r}
#| label: double-checking
 
# double checking using Spearmans
cor.test(kelp$temp_c, kelp$kelp_elong,
         method = "spearman")
```

After running the Spearman's rank test as well, I have concluded that the results lead me to the same conclusion of rejecting the null hypothesis for a moderate negative relationship between temperature and kelp frond elongation rate. Both tests have similar correlation values -0.68, along with p-values under 0.001.

# Problem 2. Personal data

## a. Updating your visualizations

## b. Captions

# Problem 3. Affective visualization

## a. Describe in words what an affective visualization could look like for your personal data

## b. Create a skietch (on paper) of your idea

## c. Make a draft of your visualization

## d. Write an artist statement

## e. Prep your materials to share in class

# Problem 4. Statistical critique

## a. Revisit and summarize

## b. Visual clarity

## c. Aesthetic clarity

## d. Recommendations



















